k-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean

### 2023-03-31 additional_files/manuscript.tex
Shorten abstract to meet GB requirements and remove dependency on csvsimple for Table 1 by embedding the table directly in the TeX file

% Packages added by M-A Lemay
\usepackage{csvsimple} % read and display csv files as tables
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables

@@

\begin{abstract} % abstract
\parttitle{Background}
Genome-wide association studies (GWAS) are powerful statistical methods that
detect associations between genotype and phenotype at genome scale. Despite
their power, GWAS frequently fail to pinpoint the causal variant or the gene
controlling a trait at a given locus in crop species. Assessing genetic
variants beyond single-nucleotide polymorphisms (SNPs) could alleviate this
problem, for example by including structural variants (SVs).

\parttitle{Results}
In this study, we tested the potential of SV- and \emph{k}-mer-based GWAS in
soybean by applying these methods to 13 traits.  We also performed
conventional GWAS analysis based on SNPs and small indels for
comparison.  We assessed the performance of each GWAS approach based on
results at loci for which the causal genes or variants were known from
previous genetic studies. We found that \emph{k}-mer-based GWAS was the
most versatile approach and the best at pinpointing causal variants or
candidate genes based on the most significantly associated
\emph{k}-mers.  Moreover, \textit{k}-mer-based analyses identified
promising candidate genes for loci related to pod color, pubescence
form, and resistance to the oomycete \textit{Phytophthora sojae}. In
our dataset, SV-based GWAS did not add value compared to
\textit{k}-mer-based GWAS and may not be worth the time and
computational resources required to genotype SVs at population scale.

@@

	\footnotesize
	\centering
	\begin{threeparttable}

		\csvreader[head to column names,
		tabular = llllccc,
		table head = \toprule Trait & Locus & Gene & Causal variant & SNPs\tnote{a} & SVs\tnote{b} & \textit{k}-mers \\ \midrule,
		table foot = \bottomrule]%
		{../tables/loci_table.csv}%
		{}%
		{\Trait & \Locus & \Gene & \Causal & \platypus & \paragraph & \kmers}

>>>>

% Packages added by M-A Lemay
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables

@@

\begin{abstract} % abstract
\parttitle{Background}
Genome-wide association studies (GWAS) are powerful statistical methods that
detect associations between genotype and phenotype at genome scale. Despite
their power, GWAS frequently fail to pinpoint the causal variant or the gene
controlling a given trait in crop species. Assessing genetic
variants beyond single-nucleotide polymorphisms (SNPs) could alleviate this
problem, for example by including structural variants (SVs).

\parttitle{Results}
In this study, we tested the potential of SV- and \emph{k}-mer-based GWAS in
soybean by applying these methods to 13 traits.  We also performed
conventional GWAS analysis based on SNPs and small indels.  We assessed the performance of each GWAS approach based on
results at loci for which the causal genes or variants were known from
previous genetic studies. We found that \emph{k}-mer-based GWAS was the
most versatile approach and the best at pinpointing causal variants or
candidate genes based on the most significantly associated
\emph{k}-mers.  Moreover, \textit{k}-mer-based analyses identified
promising candidate genes for loci related to pod color, pubescence
form, and resistance to \textit{Phytophthora sojae}. In
our dataset, SV-based GWAS did not add value compared to
\textit{k}-mer-based GWAS and may not be worth the time and
computational resources required to genotype SVs at population scale.

@@

	\footnotesize
	\centering
	\begin{threeparttable}
		\begin{tabular}{llllccc}
		\toprule Trait & Locus & Gene & Causal variant & SNPs\tnote{a} & SVs\tnote{b} & \textit{k}-mers \\ \midrule
		Flower color & \textit{W1} & Glyma.13g072100 & 65-bp deletion \citep{zabala2007} & ** & *** & *** \\
		Pubescence color & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & ** & * & ** \\
		Pubescence color & \textit{Td} & Glyma.03g258700 & SNP \citep{yan2020} & * & - & *** \\
		Seed coat color & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & ** & ** & ** \\
		Seed coat color & \textit{G} & Glyma.01g198500 & SNP \citep{wang2018} & ** & ** & *** \\
		Seed coat color\tnote{c} & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & - & - & - \\
		Seed coat color\tnote{c} & \textit{R} & Glyma.09g235100 & SNPs and indels \citep{gillman2011} & - & - & - \\
		Stem termination type & \textit{Dt1} & Glyma.19g194300 & SNPs \citep{liu2010} & ** & * & ** \\
		Stem termination type\tnote{c} & \textit{Dt2} & Glyma.18g273600 & SNPs \citep{ping2014} & - & - & - \\
		Stem termination type\tnote{c} & \textit{E3} & Glyma.19g224200 & Various types \citep{tardivel2019} & - & - & - \\
		Hilum color\tnote{c} & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & ** & ** & ** \\
		Hilum color\tnote{c} & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & * & * & ** \\
		Hilum color & \textit{R} & Glyma.09g235100 & SNPs and indels \citep{gillman2011} & * & - & ** \\
		Hilum color\tnote{c} & \textit{W1} & Glyma.13g072100 & 65-bp deletion \citep{zabala2007} & - & - & - \\
		Pubescence density & \textit{Ps} & Glyma.12g187200 & 25.6-kb CNV \citep{liu2020ps} & ** & ** & *** \\
		Pubescence density\tnote{c} & \textit{Pd1} & Glyma.01g240100 & SNP \citep{liu2020ps} & - & - & - \\
		Pubescence density\tnote{c} & \textit{P1} & Glyma.09g278000 & SNP \citep{liu2020ps} & - & - & - \\
		Seed coat luster & \textit{B} & Duplicated HPS genes & CNV (several kb) \citep{gijzen2006} & ** & ** & ** \\
		Seed coat luster\tnote{c} & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & - & - & - \\
		Maturity group\tnote{c} & \textit{E1} & Glyma.06g207800 & Various types \citep{tardivel2019} & - & - & - \\
		Maturity group\tnote{c} & \textit{E2} & Glyma.10g221500 & Various types \citep{tardivel2019} & - & - & - \\
		Maturity group\tnote{c} & \textit{E3} & Glyma.19g224200 & Various types \citep{tardivel2019} & - & - & - \\
		Maturity group\tnote{c} & \textit{E4} & Glyma.20g090000 & Various types \citep{tardivel2019} & - & - & - \\
		Oil\tnote{c} & \textit{oilGm20} & Glyma.20g085100 & 321-bp deletion \citep{fliege2022} & - & - & - \\
		Oil\tnote{c} & \textit{oilGm15} & Glyma.15g049200 & SNPs and indels \citep{zhang2020} & - & - & - \\
		Protein\tnote{c} & \textit{proGm20} & Glyma.20g085100 & 321-bp deletion \citep{fliege2022} & - & - & - \\
		Protein\tnote{c} & \textit{proGm15} & Glyma.15g049200 & SNPs and indels \citep{zhang2020} & - & - & - \\
		\bottomrule
		\end{tabular}
